---
title: "The Report of the Correlation between Healthcare Preparedness and COVID-19 Mortality"
author: "Jun, Irene, Binh, Weidun"
date: "12/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

```{r load-libraries}
library(tidyverse)
library(ggplot2)
library(magrittr)
library(dplyr)
library(readr)
library(countrycode)
library(ggthemes)
```

```{r load-csv-file}
# Load the csv file from the given source

data <- read.csv("https://wzb-ipi.github.io/corona/df_full.csv") %>%
          filter(elapsed == 334)
```

```{r process-data}
# Select columns that are desired
data <- data %>%
  select(country, continent, population_2019, pop_density, deaths_cum, deaths_cum_per_million, health_exp_pc, hosp_beds_pc, share_older, detect_index)  
```

```{r convert-country-code}
# Bring the survey dataset
# Citation: Global survey on coronavirus beliefs, behaviors, and norms: Technical Report from MIT

survey <- read.csv('survey_healthcare.csv') %>%
            select(!(gender:age)) %>%
            mutate(country = countrycode(country, 'iso2c', 'country.name'))
```

```{r merge-data}
# Bring the GHS Index
# Directly brought from the GHS Index website

ghsi <- read.csv('ghsi.csv')
data <- left_join(data, ghsi)

survey_merge <- right_join(data, survey)
```

<h3>Health Expenditure</h3>

```{r plot-health-expenses}

# Plot data to show correlation between health expenses and COVID-19 deaths per million

ggplot(data, aes(y=deaths_cum_per_million, x=health_exp_pc)) +
    theme_tufte() +
    stat_smooth(method='lm') +
    geom_point() +
    labs(title = "Health Expenses vs. COVID-19-associated deaths per million",
         subtitle = "All Countries", y = "Deaths per million", x = "Health Expenses per capita (USD)") 

```
This rather looking very unexpected, indincating more deaths for more health expenses per capita. Here are the details for the dataset.

```{r health-expenses}
 # Get correlation
cor(data$health_exp_pc, data$deaths_cum_per_million, use = "complete.obs")
 

 # Linear Regression
lm_health_exp <- lm(deaths_cum_per_million ~ health_exp_pc, data)
summary(lm_health_exp)

t.test(data$health_exp_pc, data$deaths_cum_per_million)
```

To address this unexpected pattern, we looked at one of the confounding variable: share of older population.

```{r plot-health-expenses-and-share-older}

# Add share_older to the plot as a dimension (color)

ggplot(data, aes(y=deaths_cum_per_million, x=health_exp_pc, color = share_older)) +
    theme_tufte() +
    stat_smooth(method='lm') +
    geom_point()+
    labs(title = "Health Expenses vs. COVID-19-associated deaths per million",
         subtitle = "All Countries", y = "Deaths per million", x = "Health Expenses")

```

Note that the darker-blue dots tend to stay on the left side of the graph (less health expenses), while the 
lighter-blue dots tend to stay on the right side (more health expenses). The following plot shows the correlation between the elder population share and health expenses per capita.

```{r compare-older-share-and-expenses}

# Plot between the expenditure and the share of elders

ggplot(data, aes(y=health_exp_pc, x=share_older)) +
    theme_tufte() +
    stat_smooth(method='lm') +
    geom_point()+
    labs(title = "The relationship between population share of older people and healthcare expenses",
         subtitle = "All Countries", y = "Healthcare expenses (dollars)", x = "Share of older population (%)")

 # Linear Regression
share_and_expense = lm(health_exp_pc~share_older, data)
summary(share_and_expense)
```
It gives a clear relationship between the population share and the expenditure, following the intuition that older people tend to spend more for healthcare. So, it's possible that it's not the healthcare expenses alone that might affect the mortalities but the share of older population affecting both of the healthcare expenses and the mortalities. 

We picked two groups of countries, the first having very low share of older population and another having fairly high.

```{r plot-lower-older-pop}

# Filter so that the dataset includes those with the lower share

data_control_shareolder_1 <- data %>% 
                            select(deaths_cum_per_million, share_older, health_exp_pc) %>%
                            filter(share_older < 3)

# Plot and get regression  
ggplot(data_control_shareolder_1, aes(y=deaths_cum_per_million, x=health_exp_pc, color = share_older)) +
    theme_tufte() +
    stat_smooth(method='lm') +
    geom_point() +
    labs(title = "The relationship between healthcare expenses and mortality rate",
         subtitle = "For those countries having low share_older (less than 3 percent)",
         y = "Deaths per million", x = "Healthcare expenditure (dollars)")

lm_control_shareolder_1 <- lm(deaths_cum_per_million ~ health_exp_pc, data = data_control_shareolder_1)
summary(lm_control_shareolder_1)
```
Due to the cluster at the lowest healthcare expenditure, The corrlation becomes highly significant, although the pattern did not result yet again not as expected.

Despite the more clear pattern for the countries with the lower-share, the pattern becomes very unsignificant the countries with the higher-share.

```{r plot-higher-older-pop}

# Filter so that the dataset will include those with the higher share only
data_control_shareolder_2 <- data %>% 
                            select(deaths_cum_per_million, share_older, health_exp_pc) %>%
                            filter(share_older > 10)

# Plot and regression  
ggplot(data_control_shareolder_2, aes(y=deaths_cum_per_million, x=health_exp_pc, color = share_older)) +
    theme_tufte() +
    stat_smooth(method='lm') +
    geom_point() +
    labs(title = "The relationship between healthcare expenses and mortality rate",
         subtitle = "For those countries having low share_older (less than 3 percent)",
         y = "Deaths per million", x = "Healthcare expenditure (dollars)")

lm_control_shareolder_2 <- lm(deaths_cum_per_million ~ health_exp_pc, data = data_control_shareolder_2)
summary(lm_control_shareolder_2)
```
Thus, in spite of the control, Healthcare expenditure and deaths per million have positive correlation.


Number of hospital beds simply means the capacities of hospitals to admit patients. We looked into the correlation between the number of hospital beds (per capita) and the mortalities.

```{r plot-hosptal-beds}

# Plot between hospital beds per capita and mortality
ggplot(data, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "All countries", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')


 # Get confidence interval
hosp_beds_and_mortality = lm(deaths_cum_per_million~hosp_beds_pc, data)
summary(hosp_beds_and_mortality)
```
It merely shows a weak correlation with r-squared number around 0.04.

However, when we introduce regions, the 
```{r plot-hospital-bed-with-regions}

# Same plot but a dimension added: continents
ggplot(data, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million, color = continent)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "All countries", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')
```
Each region shows different patterns, with some having as expected.

We dived groups by continents, and performed linear regression on each group.

```{r divide-groups-by-region}
data_control_region <- data %>%
                        select(hosp_beds_pc, continent, deaths_cum_per_million)

# Divide groups by regions
data_control_region_america <- data_control_region %>% filter(continent == 'America')

data_control_region_asia <- data_control_region %>% filter(continent == 'Asia')

data_control_region_europe <- data_control_region %>% filter(continent == 'Europe')

data_control_region_africa <- data_control_region %>% filter(continent == 'Africa')

data_control_region_oceania <- data_control_region %>% filter(continent == 'Oceania')
```

```{r plot-america}
ggplot(data_control_region_america, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "Americas", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')

 # Get confidence interval
lm_hosp_bed_am = lm(deaths_cum_per_million~hosp_beds_pc, data_control_region_america)
summary(lm_hosp_bed_am)
```

```{r plot-europe}
ggplot(data_control_region_europe, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "Europe", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')

 # Get confidence interval
lm_hosp_bed_eu = lm(deaths_cum_per_million~hosp_beds_pc, data_control_region_europe)
summary(lm_hosp_bed_eu)
```
```{r plot-asia}
ggplot(data_control_region_asia, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "Asia", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')

 # Get confidence interval
lm_hosp_bed_as = lm(deaths_cum_per_million~hosp_beds_pc, data_control_region_asia)
summary(lm_hosp_bed_as)
```
```{r plot-oceania}
ggplot(data_control_region_oceania, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "Oceania", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')

 # Get confidence interval
lm_hosp_bed_oc = lm(deaths_cum_per_million~hosp_beds_pc, data_control_region_oceania)
summary(lm_hosp_bed_oc)
```

```{r plot-africa}
ggplot(data_control_region_africa, mapping = aes(x = hosp_beds_pc, y = deaths_cum_per_million)) +
  geom_point() +
  theme_tufte() +
  labs(title = "COVID-19 mortality rate and number of hospital beds available", 
       subtitle = "Africa", 
       x = "Number of hospital beds", y = "Deaths per Million") +
  stat_smooth(method='lm')

 # Get confidence interval
lm_hosp_bed_af = lm(deaths_cum_per_million~hosp_beds_pc, data_control_region_africa)
summary(lm_hosp_bed_af)
```

```{r plot-ghsi}

# Plot between GHS Index and mortality
ggplot(data, mapping = aes(x = ghsi, y = deaths_cum_per_million)) +
  geom_point() +
  labs(title = "COVID-19 mortality rate and health index", subtitle = "All countries", 
       x = "Health index", y = "Deaths per million") +
  stat_smooth(method='lm') +
  theme_tufte()
```
```{r ghsi-regression}
# Get the correlation coefficient
# cor(data_health_index$health_index, data_health_index$deaths_cum_per_million, use = "complete.obs")

 # Get confidence interval
ghsi_and_mortality = lm(deaths_cum_per_million~ghsi, data)
summary(ghsi_and_mortality) 

t.test(data$ghsi, data$deaths_cum_per_million)
```
It shows a positive correlation, with a weak r-squared of 0.157 but plausible p-value.

We then look at the relationship between detection and GHS Index and that of detection and the mortalities. We brought the detect_index and compared to GHS index:

```{r plot-detection-ghsi}

# Plot between the detect_index and GHS Index
ggplot(data, mapping = aes(x = detect_index, y = ghsi)) +
  geom_point() +
  labs(title = "COVID-19 mortality rate and health index", subtitle = "All countries", 
       x = "Detect Index", y = "Deaths per million") +
  stat_smooth(method='lm') +
  theme_tufte()

lm_detect_and_ghsi = lm(ghsi~detect_index, data)
summary(lm_detect_and_ghsi) 
```
Since detect index is one of the measurement from the GHS Index, The correlation is strong and reliable with r-squared = 0.775 and extremely low p-value.


```{r plot-detection}

# Plot detection and mortality; add GHS Index as another dimension
ggplot(data, mapping = aes(x = detect_index, y = deaths_cum_per_million, color = ghsi)) +
  geom_point() +
  labs(title = "COVID-19 mortality rate and health index", subtitle = "All countries", 
       x = "Detect Index", y = "Deaths per million") +
  stat_smooth(method='lm') +
  theme_tufte()

detect_and_mortality = lm(deaths_cum_per_million~detect_index, data)
summary(detect_and_mortality) 
```
Here, the detection index and the mortality rate are weakly correlative.

